### 2010-01-22 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Final version before Albert's input
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@451 d8b0c59b-6544-0410-94d9-83a5658a058e

\paragraph{Structure of the Document.}
Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment. Section \ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand. Thus, Section \ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilites of the tool, and the testing procedure with the results.  The remaing part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue. Section \ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Section \ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.

%Describe the rest.
Part \ref{part:d2.4.1} contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching. We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology mediation problem in Section \ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in Section \ref{sec:mediationapproaches}.

>>>>

\paragraph{Structure of the Document.}
Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment. Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand. Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilites of the tool, and the testing procedure with the results.  The remaing part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue. Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.

%Describe the rest.
Part \ref{part:d2.4.1} contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching. We discuss ontologies in general in Sect.~\ref{sec:ontologies}, present the ontology mediation problem in Sect.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in Sect.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in Sect.~\ref{sec:mediationapproaches}.

### 2010-01-22 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Final version before Albert's input
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@451 d8b0c59b-6544-0410-94d9-83a5658a058e

\begin{inparaenum}[(a)]
    \item A \textit{level 0} mapping~\cite{euzenat2004api} is a set of the above mapping elements, when the entities are discreet (defined by URIs). E.g., consider the ontology $O1$ with a class \texttt{Person}, and another ontology $O2$ with a class \texttt{Human}. For this case a matching algorithm  could return the mapping element $< id_{11}, Person, Human, 0.67, = >$, meaning that the \texttt{Person} class from the first ontology is found to be equivalent to the \texttt{Human} class in the second one with a confidence measure of \texttt{0.67}.
    \item A \textit{level 1} mapping is a slight refinement of level 0, replacing pairs of elements with pairs of sets of elements.
    \item A \textit{level 2} mapping can be more complex and defines correspondences in first order logic. It uses the ontology mapping language described in~\cite{scharffe2005language}. It can describe complex correspondences, such as the one detailed in Section \ref{sub:manual_mapping}.
\end{inparaenum}
% :
% 
% \begin{equation*} \label{eq}
% \forall{x,z} grandparent(x,z) \Longrightarrow \exists{y}; parent(x,y) \wedge parent(y,z)
% \end{equation*}

@@

The gadget life cycle in FAST has several phases and roles associated, as detailed in \cite{hoyer-fast}. Here, we list the ones relevant for the ontology matching tasks, in decreasing order of the measure in which knowledge about ontologies is required. Note that several roles can be played by the same actor.
\begin{enumerate}[(i)]
    \item The \textit{ontology engineer} creates the ontologies used to annotate services and data. This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST screen building. The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions.
    \item Ontology matching is needed by the \textit{gadget developer} at the design-time of a screen-flow (gadget). Gadget developers use the GVS (Gadget Visual Storyboard) for building gadgets, in which they can use the matching operators to combine screens having input and output resources annotated with different ontologies. No actual matching needs to be performed in this phase, but rather the possibility of matching needs to be determined (i.e., \emph{can} two screens A and B be combined?).
%the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
    %\item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
    \item The \textit{end-user} uses the final deployed gadget at run-time, but is unaware of the underlying resources and ontologies or the matching process. Only at run-time the actual mapping of instance data has to be performed. In this paper, we mainly consider the first two cases (i.e. ontology engineering and screen development).
\end{enumerate}

@@

Based on these requirements, we compared three different tools. \emph{MAFRA}~\cite{maedche2002mafra} supports an interactive and incremental process of ontology mapping. It provides an explicit notion of semantic bridges. This representation is serialisable, portable and independent from the mapped languages. The bridges, however, have been designed to be used within the MAFRA system, and the alignment process needs to be done through the provided GUI. \emph{RDFT}~\cite{omelayenko2002rdft} is a small language originally designed to map between XML and RDF. The results are mappings represented in DAML+OIL, that can be executed in a transformation process. No hints are given to add alignment methods or extending the format and the tool does not longer seem to be available. \emph{Alignment API}~\cite{euzenat2004api} is the tool best matching our requirements, satisfying all the desired conditions. It is still under active development, provides an API and its implementation, is open source (GPLv2 or above) and written in Java, providing an easy way to embed it into other programs. Alignment API can be extended by other representations and matching algorithms, it can be invoked through the command line interface (thus working without user interference) or one of the two available GUI implementations, or it can be exposed as an HTTP server. The tool allows for testing different alignment methods and can generate evaluation results based on a reference alignment. Alignment API can generate the mapping results in XSLT, therefore providing an easy way to integrate them into other systems.

@@

Though the infeasibility of automating this alignment became obvious, we have represented the alignment in the mapping language supported by the tool, as a level 2 mapping (described in Sect. \ref{mediation}).
 % (the mapping has been created between Amazon and GoodRelations, since the eBay and Amazon ontologies are very similar in approach, and mapping eBay to GoodRelations would be an indentical task). 
This mapping description can later be used by the run-time gadget code. List.~\ref{listing_gr_a} shows an example mapping between two properties of the two ontologies, specifying that the relationship is \texttt{Equivalence} with a certainty degree of \texttt{1.0}. This fragment does not show, but assumes the equivalence correspondence between the classes \texttt{Item} and \texttt{Offering}, which is a trivial level 0 mapping. This mapping specifies the relation

>>>>

\begin{inparaenum}[(a)]
    \item A \textit{level 0} mapping~\cite{euzenat2004api} is a set of the above mapping elements, when the entities are discreet (defined by URIs). E.g., consider the ontology $O1$ with a class \texttt{Person}, and another ontology $O2$ with a class \texttt{Human}. For this case a matching algorithm  could return the mapping element $< id_{11}, Person, Human, 0.67, = >$, meaning that the \texttt{Person} class from the first ontology is found to be equivalent to the \texttt{Human} class in the second one with a confidence measure of \texttt{0.67}.
    \item A \textit{level 1} mapping is a slight refinement of level 0, replacing pairs of elements with pairs of sets of elements.
    \item A \textit{level 2} mapping can be more complex and defines correspondences in first order logic. It uses the ontology mapping language described in~\cite{scharffe2005language}. It can describe complex correspondences, such as the one detailed in Sect.~\ref{sub:manual_mapping}.
\end{inparaenum}
% :
% 
% \begin{equation*} \label{eq}
% \forall{x,z} grandparent(x,z) \Longrightarrow \exists{y}; parent(x,y) \wedge parent(y,z)
% \end{equation*}

@@

The gadget life cycle in FAST has several phases and roles associated, as detailed in \cite{hoyer-fast}. Here, we list the ones relevant for the ontology matching tasks, in decreasing order of the measure in which knowledge about ontologies is required. Note that several roles can be played by the same actor.
\begin{enumerate}[(i)]
    \item The \textit{ontology engineer} creates the ontologies used to annotate services and data. This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST gadget building. The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions of a screen.
    \item Ontology matching is needed by the \textit{gadget developer} at the design-time of a screen-flow (gadget). Gadget developers use the GVS (Gadget Visual Storyboard) for building gadgets, where they can use the matching operators to combine screens having input and output conditions specifying resources annotated with different ontologies. No actual matching needs to be performed in this phase, but rather the possibility of matching needs to be determined (i.e., \emph{can} two screens A and B be combined?).
%the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
    %\item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
    \item The \textit{end-user} uses the final deployed gadget at run-time, but is unaware of the underlying resources and ontologies or the matching process. Only at run-time the actual mapping of instance data has to be performed.
\end{enumerate}

@@

Based on these requirements, we compared three different tools. 
\paragraph{MAFRA} \emph{MAFRA}~\cite{maedche2002mafra} supports an interactive and incremental process of ontology mapping. It provides an explicit notion of semantic bridges. This representation is serialisable, portable and independent from the mapped languages. The bridges, however, have been designed to be used within the MAFRA system, and the alignment process needs to be done through the provided GUI.

\paragraph{RDFT} \emph{RDFT}~\cite{omelayenko2002rdft} is a small language originally designed to map between XML and RDF. The results are mappings represented in DAML+OIL, that can be executed in a transformation process. No hints are given to add alignment methods or extending the format and the tool does not longer seem to be available. 
\paragraph{Alignment API} \emph{Alignment API}~\cite{euzenat2004api} is the tool best matching our requirements, satisfying all the desired conditions. It is still under active development, provides an API and its implementation, is open source (GPLv2 or above) and written in Java, providing an easy way to embed it into other programs. Alignment API can be extended by other representations and matching algorithms, it can be invoked through the command line interface (thus working without user interference) or one of the two available GUI implementations, or it can be exposed as an HTTP server. The tool allows for testing different alignment methods and can generate evaluation results based on a reference alignment. Alignment API can generate the mapping results in XSLT, therefore providing an easy way to integrate them into other systems.

@@

Though the infeasibility of automating this alignment became obvious, we have represented the alignment in the mapping language supported by the tool, as a level 2 mapping (described in Sect.~\ref{mediation}).
 % (the mapping has been created between Amazon and GoodRelations, since the eBay and Amazon ontologies are very similar in approach, and mapping eBay to GoodRelations would be an indentical task). 
This mapping description can later be used by the run-time gadget code. List.~\ref{listing_gr_a} shows an example mapping between two properties of the two ontologies, specifying that the relationship is \texttt{Equivalence} with a certainty degree of \texttt{1.0}. This fragment does not show, but assumes the equivalence correspondence between the classes \texttt{Item} and \texttt{Offering}, which is a trivial level 0 mapping. This mapping specifies the relation
